% Created 2019-08-20 Tue 16:50
% Intended LaTeX compiler: pdflatex
\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Jonathan Tow}
\date{\today}
\title{}
\hypersetup{
 pdfauthor={Jonathan Tow},
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.2 (Org mode 9.1.9)}, 
 pdflang={English}}
\begin{document}

\textbf{Assignment 3 - Written}: Machine Learning \& Neural Networks


\section*{1. Machine Learning \& Neural Networks}
\label{sec:org864511c}

\textbf{(a)} Adam Optimizer.

(i) Adam optimization uses a trick called \emph{momentum} by keeping track of \textbf{m}, a rolling average of the gradients:

\begin{LATEX}
\begin{align*}
\bold{m_{i}} &\leftarrow \beta_1 \bold{m_{i-1}} + (1-\beta_1) \nabla_{\theta} J_{minibatch}(\theta) \\
\theta &\leftarrow \theta - \alpha\bold{m_i}
\end{align*}
\end{LATEX}

where \(\beta_1\) is a hyperparameter between 0 and 1. Briefly explain how using \textbf{m} stops the update from varying 
as much and why this low variance may be helpful to learning, overall.\\


\textbf{Solution}: By maintaining a sort of exponential smoothing, or rolling average, of the loss function's gradients, we effectively control 
a "memory bank" that tracks previous update steps and fuses them together to some degree, \(\beta_1\), with new gradient information. This can
smooth transitions, i.e. lower variance, from one state to the next when we place the right amount of \(\beta_1\) on the previous state
\(\bold{m}_{i-1}\) so that recent changes have a bit more importance than the newly retrieved gradient at position \(\theta\). Note that when we don't
care about  the past, \(\beta_1 = 0\), we get back simple SGD.
This dampening of oscillations is helpful because it straightens out the trail that descends into the minimum. 
Informally, the efficiency of optimization increases when update steps zig-zag (vary) less.\\

(ii) Adam also uses \emph{adaptive learning rates} by keeping track of \(\bold{v}\), a rolling average of the magnitudes of the gradients:

\begin{LATEX}
\begin{align*}
\bold{m} &\leftarrow \beta_1 \bold{m} + (1-\beta_1) \nabla_{\theta} J_{minibatch}(\theta) \\
\bold{v} &\leftarrow \beta_2 \bold{v} + (1-\beta_2) (\nabla_{\theta} J_{minibatch}(\theta) \odot \nabla_{\theta} J_{minibatch}(\theta))\\
\theta &\leftarrow \theta - \alpha \odot \bold{m} / \sqrt{\bold{v}}
\end{align*}
\end{LATEX}

Since Adam divides the update by \(\sqrt{\bold{v}}\) which of the model parameters will get larger updates? Why might this help 
learning?\\

\textbf{Solution:} Because \(\bold{v}\) is a rolling average of the magnitudes of gradients, parameters of the model 
with corresponding small gradients, i.e. small \(\bold{v}\) entries, will get larger updates. We're cutting out
a smaller amount from such parameters during an update because we divide them by their small gradient 
magnitude/norm. 
This might be helpful because it can get recent stagnant parameters moving more efficiently along their axes 
and in turn expedite convergence.\\

\textbf{(b)} Dropout is a regularization technique. During training, dropout randomly sets units in the hidden layer \(\bold{h}\) to
 zero with probability \(p_{drop}\) (dropping different units in each minibatch), and then multiplies \(\bold{h}\) by a constant
 \(\gamma\). We can write this as:

\[\bold{h}_{drop} = \gamma \bold{d} \circ \bold{h}\]

where \(\bold{d} \in \{0, 1\}^{D_k}\) is a mask vector.\\

(i) What must \(\gamma\) equal in terms of \(p_{drop}\)?\\

\textbf{Solution:}
During training we drop units at a rate of \(p_{drop}\), resulting in roughly \(p_{keep} = 1 - p_{drop}\) fraction of
units left over. At test time we'd like to have the effect of keeping a similar fraction,   '\(p_{keep}\), of 
units on. By scaling down the layer units by \(\gamma = \frac{1}{1-p_{drop}} = \frac{1}{p_{keep}}\), we effectively level out their magnitudes so that
both phases of learning share very similar expected outputs.\\

(ii) Why should we apply dropout during training but not during evaluation?\\

\textbf{Solution:}
The goal of dropout is to reduce overfitting. We're interested in updating unit weights so as to form a network that
performs well across different datasets. Now, during evaluation we're concerned with how well the model handles unseen
data. When we dropout units, we're "thinning" out the network which in many cases will add noise to predictions and 
dampen accuracy. Thus, if we were to apply dropout during evaluation time, we would not be able to fairly assess the
generalization power of the network.

\newpage

\section*{2. Neural Transition-Based Dependency Parsing}
\label{sec:orgf1df58f}

\textbf{(a)} \textbf{Transition-Based Parse}: A parser which incrementally builds up a parse one step at a time. At every step it 
maintains a \emph{partial parse} which is represented as:

\begin{itemize}
\item A \emph{stack} of words that are currently being processed.
\item A \emph{buffer} of words yet to be processed.
\item A list of \emph{dependencies} predicted by the parser.
\end{itemize}

Initially the stack contains ROOT, the dependencies list is empty, and the buffer contains all words of the sentence 
in order. At each step the parser applies a \emph{transition} to the partial parse until its buffer is empty and the stack 
size is 1. The following transitions can be applied:

\begin{itemize}
\item \emph{SHIFT}: removes the first word from the buffer and pushes it onto the stack.
\item \emph{LEFT-ARC}: marks the second (second most recently added) item on the stack as a dependent of the first item and 
removes the second item from the stack.
\item \emph{RIGHT-ARC}: marks the first (most recently added) item on the stack as a dependent of the second item and removes
the first item from the stack.\\
\end{itemize}

\textbf{Solution}:

\begin{longtable}{|l|l|l|l|}
Stack & Buffer & New Dependency & Transition\\
\hline
\endfirsthead
\multicolumn{4}{l}{Continued from previous page} \\
\hline

Stack & Buffer & New Dependency & Transition \\

\hline
\endhead
\hline\multicolumn{4}{r}{Continued on next page} \\
\endfoot
\endlastfoot
\hline
(ROOT) & [I, parsed, this, sentence, correctly] &  & Initial Config\\
(ROOT, I) & [parsed, this, sentence, correctly] &  & SHIFT\\
(ROOT, I, parsed) & [this, sentence, correctly] &  & SHIFT\\
(ROOT, parsed) & [this, sentence, correctly] & parsed->I & LEFT-ARC\\
(ROOT, parsed, this) & [sentence, correctly] &  & SHIFT\\
(ROOT, parsed, this, sentence) & [correctly] &  & SHIFT\\
(ROOT, parsed, sentence) & [correctly] & sentence->this & LEFT-ARC\\
(ROOT, parsed) & [correctly] & parsed->sentence & RIGHT-ARC\\
(ROOT, parsed, correctly) & [] &  & SHIFT\\
(ROOT, parsed) & [] & parsed->correctly & RIGHT-ARC\\
(ROOT) & [] & root->parsed & RIGHT-ARC\\
\end{longtable}

\textbf{(b)} How many steps will it take to parse \(n\) words (in terms of \(n\))?

\textbf{Solution}: In the worst case, parsing will take linear time, i.e. \(O(n)\). At any step of parsing, we have two possible
state transitions, either shifting a word from the buffer to the stack or clearing a dependent from the stack. Every 
word must spend a single step being shifted from the buffer, thus \(n\) words cost \(n\) shift steps. From the stack a 
word must be "arc"-ed over as a dependent exactly once, thus \(n\) words cost \(n\) "arc"-ing steps. Therefore, we have 
\(2*n\) steps giving a cost of \(O(n)\).


\textbf{(e)} Report of best UAS model:

\begin{longtable}{|c|c|}
dev UAS & test UAS\\
\hline
\endfirsthead
\multicolumn{2}{l}{Continued from previous page} \\
\hline

dev UAS & test UAS \\

\hline
\endhead
\hline\multicolumn{2}{r}{Continued on next page} \\
\endfoot
\endlastfoot
\hline
89.60 & 89.74\\
\end{longtable}

\textbf{(f)} For each sentence state the type of error, the incorrect dependency, and the correct dependency:\\

(i)\\

\begin{itemize}
\item \textbf{Error Type}:
\item \textbf{Incorrect Dependency}:
\item \textbf{Correct Dependency}:
\end{itemize}

(ii)\\

\begin{itemize}
\item \textbf{Error Type}:
\item \textbf{Incorrect Dependency}:
\item \textbf{Correct Dependency}:
\end{itemize}

(iii)\\

\begin{itemize}
\item \textbf{Error Type}:
\item \textbf{Incorrect Dependency}:
\item \textbf{Correct Dependency}:
\end{itemize}

(iv)\\

\begin{itemize}
\item \textbf{Error Type}:
\item \textbf{Incorrect Dependency}:
\item \textbf{Correct Dependency}:
\end{itemize}
\end{document}
